---
title: "kernel_ml_fellipe"
author: "MTBR - ModelThinkingBR"
date: "4/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE, result=FALSE, cache=FALSE)
knitr::knit_hooks$set(
  source = function(x, options) {
    hook.r = function(x, options) {
      fence <- '```'
      language = tolower(options$engine)
      if (language == 'node') language = 'javascript'
      if (!options$highlight) language = 'text'
      if(!is.null(options$foldcode)) {
      paste0('\n\n', "<details><summary>Source</summary>\n", fence, language, '\n', x, fence,  '\n\n', "</details>\n")
      } else {
              paste0('\n\n', fence, language, '\n', x, fence,  '\n\n')
      }
    }
    x = knitr:::hilight_source(x, 'markdown', options)
    hook.r(
      paste(c(
        x, 
        ''
      ), collapse = '\n'), 
      options
    )
  }
)
```

Project authors and participants:

* [Fellipe Gomes](https://github.com/gomesfellipe) (Statistics - UFF, Data Scientist - FGV IBRE / BRAZIL)
* [Rumenick Pereira da Silva](https://github.com/Rumenick) (PhD Statistics - UFMG, Data Scientist - FGV IBRE / BRAZIL)
* [Luiz Fernando Coelho Passos](https://github.com/luizfcp) (Statistics student - UFF, Data Scientist Intern - FGV IBRE / BRAZIL)


<p align="right"><span style="color:firebrick">If you liked the kernel dont forget the upvote! <i class="fas fa-hand-peace"></i></span> </p>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# packages
library(readr)      # read/write
library(dplyr)      # manipulate data
library(tidyr)      # tidy data
library(purrr)      # functional programming
library(stringr)    # text manipulation
library(quanteda)   # textmining
library(knitr)      # print table
library(kableExtra) # custom table
library(ggplot2)    # elegant plot
library(GGally)     # Pairs
library(gghalves)   # half plot
theme_set(theme_bw()) # set theme

# getting data:
train_data <- read_csv("../data/train.csv")
source("../functions.R")
metadata <- get_metadata(train_data)

```


